
---
title: "Lab1"
output:
  html_document:
    df_print: paged
---

              LAB 1
              
              
              
Exercise  1.1 INPUT: 60 85 72 59 37 75 93 7 98 63 41 90 5 17 97
```{r}
x=c(60,85,72,59,37,75,93,7,98,63,41,90,5,17,97)
sx=stem(x)
print(sx)
```
Exercise 1.2
#########################################################
 The decimal point is 1 digit(s) to the right of the |

   8 | 028
   9 | 115578
  10 | 1669
  11 | 01   
##########################################################

The data from the given stem and leaf plot are: 80,82,88,91,91,95,95,97,98,101,106,106,109,110,111
```{r}
data=c(80,82,88,91,91,95,95,97,98,101,106,106,109,110,111)
stem(data)
```
```{r}
hist(data,probability=TRUE)
```
Exercise 1.3
1st Randomization
```{r}
f=rnorm(100)
hist(f)
```
2nd Randomization
```{r}
f=rnorm(100)
hist(f)
```
As we see the two histograms are not the same


Exercise 1.4

library("UsingR")
```{r}
library("UsingR")
```
For south dataset:
```{r}
data(south)
hist(south)
```
```{r}
boxplot(south,horizontal = TRUE)
```

For crime:
```{r}
data(crime)
hist(crime)
```
```{r}
boxplot(crime,horizontal = TRUE)
```
For aid dataset:
```{r}
data(aid)
hist(aid)
```
```{r}
boxplot(aid,horizontal=TRUE)
```
As we see from the diagrams above the south dataset has the most skewed data, in addition to the crime dataset which data are much less skewed.The only symmetry we can find is in aid dataset.

Exercise 1.5

a) Predict for bumpers:
```{r}
data(bumpers)
hist(bumpers)
```
Predictions:
  Median:2100
  Mean:2000
  Standard deviation:500
  
```{r}
median(bumpers)
mean(bumpers)
sd(bumpers)
```
b) Predict for firstchi:
```{r}
data(firstchi)
hist(firstchi)
```
Predictions:
  Median:26
  Mean:27.5
  Standard deviation:5
  
```{r}
median(firstchi)
mean(firstchi)
sd(firstchi)
```
c) Predict for math:
```{r}
data(math)
hist(math)
```

Predictions:
  Median:55
  Mean:55
  Standard deviation:10
  
```{r}
median(math)
mean(math)
sd(math)
```
Exercise 1.6
 Number of O-ring failures for the first 23 flights of the US space shuttle Challenger : 0,1,0,NA,0,0,0,0,0,1,1,1,0,0,3,0,0,0,0,0,2,0,1

```{r}
flights=c(0,1,0,NA,0,0,0,0,0,1,1,1,0,0,3,0,0,0,0,0,2,0,1)
fl=table(flights)
print(fl)
mean(fl[!is.na(fl)])
```
Exercise 1.7

```{r}
hist(pi2000,probability = TRUE)
```
```{r}
prop=table(pi2000)/length(pi2000)
print(prop)
print(prop[2])
print(prop[3])
print(prop[4])
```
We can estimate the proportions for all digits
The proportion for 1's,2's and 3's are:
1)0.1065
2)0.1035
3)0.0945

Exercise 1.8
Fit a density to pi2000

```{r}
data(pi2000)
#attach(prop)
hist(pi2000,probability=TRUE)
lines(density(pi2000,bw=0.45))
```

























