---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

                            Lab 2
                            
        RANDOM DATA
        
          Although Einstein said that god does not play dice, R can. For example
          
```{r}
               sample(1:6,10,replace=T)
```
          or with a function
          
```{r}
          RollDie = function(n) sample(1:6,n,replace=T)
          RollDie(5)
```
  2.1  Random number generators in R -  the "r" functions.
          
          
      As we know, random numbers are described by a distribution. That is, some       function which specifies the probability that a random number is in some        range. For example P(a<X<??????b). Often this is given by a probability           density (in the continuous case) or by a function P(X=k) = f(k) in the          discrete case.R will give numbers drawn from lots of different                 distributions. In order to use them, you only need familiarize yourselves       with the parameters that are given to the functions such as a mean, or a        rate. Here are examples of the most common ones. For each, a histogram is       given for a random sample of size 100, and density (using the d''               functions) is superimposed as appropriate.
      Uniform.
      Uniform numbers are ones that are "equally likely" to be in the specified       range. Often these numbers are in [0,1] for computers, but in practice can       be between [a,b] where a,b depend upon the problem. An example might be         the time you wait at a traffic light. This might be uniform on [0,2].
      
      
```{r}
        runif(1,0,2)                  # time at light
        #also runif(1,min=0,max=2)

```
```{r}
      runif(5,0,2)                  # time at 5 lights
```
```{r}
    runif(5)                      # 5 random numbers in [0,1]
```
    The general form is runif(n,min=0,max=1) which allows you to decide how many     uniform random numbers you want (n), and the range they are chosen from         ([min,max])

    To see the distribution with min=0 and max=1 (the default) we have
    
```{r}
    x=runif(100)                 # get the random numbers
    hist(x,probability=TRUE,col=gray(.9),main="uniform on [0,1]")    
    curve(dunif(x,0,1),add=T)
```
    
        The only tricky thing was plotting the histogram with a                         background color''. Notice how the dunif function was used with                 the curvefunction.
  Normal.
        Normal numbers are the backbone of classical statistical theory due to          the central limit theorem The normal distribution has two parameters a          mean ?? and a standard deviation ???. These are the location and spread            parameters. For example, IQs may be normally distributed with mean 100          and standard deviation 16, Human gestation may be normal with mean 280          and standard deviation about 10 (approximately). The family of normals          can be standardized to normal with mean 0 (centered) and variance 1.            This is achieved by "standardizing" the numbers, i.e. 
        Z=(X- ??)/?????.
        
        Here are some examples
```{r}
        rnorm(1,100,16)               # an IQ score
```
```{r}
      rnorm(1,mean=280,sd=10)       # how long for a baby (10 days early) 
```
        Here the function is called as rnorm(n,mean=0,sd=1) where one specifies         the mean and the standard deviation.

        To see the shape for the defaults (mean 0, standard deviation 1) we have
```{r}
          x=rnorm(100)
          hist(x,probability=TRUE,col=gray(.9),main="normal mu=0,sigma=1")
          curve(dnorm(x),add=T)
          ## also for IQs using rnorm(100,mean=100,sd=16)
```
```{r}
          x=rnorm(100,mean=100,sd=16)
          hist(x,probability=TRUE,col=gray(.9),main="normal mean=100,sd=16")
          curve(dnorm(x,mean=100,sd=16),add=T)  
```
        
                 
                 
                 
                 
                 
                 
                      Binomial.
	
1) Compute probability according to binomial distribution
The binomial distribution is a discrete probability distribution. It describes the outcome of n independent trials in an experiment. Each trial is assumed to have only two outcomes, either success or failure. If the probability of a successful trial is p, then the probability of having x successful outcomes in an experiment of n independent trials is as follows.

Problem
Suppose there are twelve multiple choice questions in an English class quiz. Each question has five possible answers, and only one of them is correct. Find the probability of having four or less correct answers if a student attempts to answer every question at random.
Solution
Since only one out of five possible answers is correct, the probability of answering a question correctly by random is 1/5=0.2. We can find the probability of having exactly 4 correct answers by random attempts as follows.

```{r}
    dbinom(4,size=12,prob=0.2)
```
To find the probability of having four or less correct answers by random attempts, we apply the function dbinom with x = 0,.,4.
```{r}
    dbinom(0,size=12,prob=0.2)+
dbinom(1,size=12,prob=0.2)+
dbinom(2,size=12,prob=0.2)+
dbinom(3,size=12,prob=0.2)+
dbinom(4,size=12,prob=0.2)
```
Alternatively, we can use the cumulative probability function for binomial distribution pbinom.
```{r}
      pbinom(4,size=12,prob=0.2)
```
2) Generate binomial random numbers
The binomial random numbers are discrete random numbers. They have the distribution of the number of successes in n independent Bernoulli trials where a Bernoulli trial results in success or failure, success with probability p.

A single Bernoulli trial is given with n=1 in the binomial
```{r}
      n=1
      p=.5                     # set the probability
      rbinom(1,n,p)                 # different each time
```
```{r}
        rbinom(10,n,p)                # 10 different such numbers
```
A binomially distributed number is the same as the number of 1's in n such Bernoulli numbers. For the last example, this would be 5. There are then two parameters n (the number of Bernoulli trials) and p (the success probability).

To generate binomial numbers, we simply change the value of n from 1 to the desired number of trials. For example, with 10 trials:
```{r}
      n = 10; p=.5
      rbinom(1,n,p)                 # 6 successes in 10 trials
```
```{r}
      rbinom(5,n,p)                 # 5 binomial number
```

The graphs (figure 27) show 100 binomially distributed random numbers for 3 values of n and for p=.25. Notice in the graph, as n increases the shape becomes more and more bell-shaped. These graphs were made with the commands
```{r}
      n=5;p=.25                    # change as appropriate
      x=rbinom(100,n,p)            # 100 random numbers
     hist(x,probability=TRUE)
## use points, not curve as dbinom wants integers only for x
     xvals=0:n;points(xvals,dbinom(xvals,n,p),type="h",lwd=3) 
     points(xvals,dbinom(xvals,n,p),type="p",lwd=3) 
##...    repeat with n=15, n=50
```
```{r}
  n=15;p=.25                    # change as appropriate
      x=rbinom(100,n,p)            # 100 random numbers
     hist(x,probability=TRUE)
## use points, not curve as dbinom wants integers only for x
     xvals=0:n;points(xvals,dbinom(xvals,n,p),type="h",lwd=3) 
     points(xvals,dbinom(xvals,n,p),type="p",lwd=3)
```
```{r}
  n=50;p=.25                    # change as appropriate
      x=rbinom(100,n,p)            # 100 random numbers
     hist(x,probability=TRUE)
## use points, not curve as dbinom wants integers only for x
     xvals=0:n;points(xvals,dbinom(xvals,n,p),type="h",lwd=3) 
     points(xvals,dbinom(xvals,n,p),type="p",lwd=3)
```

Exponential
The exponential distribution is important for theoretical work. It is used to describe lifetimes of electrical components (to first order). For example, if the mean life of a light bulb is 2500 hours one may think its lifetime is random with exponential distribution having mean 2500. The one parameter is the rate = 1/mean. We specify it as follows rexp(n,rate=1/2500). Here is an example with the rate being 1/2500
```{r}
      x=rexp(100,1/2500)
     hist(x,probability=TRUE,col=gray(.9),main="exponential mean=2500")
     curve(dexp(x,1/2500),add=T)
```
There are others of interest in statistics. Common ones are the Poisson, the Student t-distribution, the F distribution, the beta distribution and the ???2 (chi squared) distribution.


2.2  Sampling with and without replacement using sample

R has the ability to sample with and without replacement. That is, choose at random from a collection of things such as the numbers 1 through 6 in the dice rolling example. The sampling can be done with replacement (like dice rolling) or without replacement (like a lottery). By default sample samples without replacement each object having equal chance of being picked. You need to specify replace=TRUE if you want to sample with replacement. Furthermore, you can specify separate probabilities for each if desired.

Here are some examples 

```{r}
## Roll a die
 sample(1:6,10,replace=TRUE)
```
# no sixes!
```{r}
## toss a coin
 sample(c("H","T"),10,replace=TRUE)
```
```{r}
## pick 6 of 54 (a lottery)
 sample(1:54,6)                # no replacement
```
```{r}
## pick a card. (Fancy! Uses paste, rep)
cards = paste(rep(c("A",2:10,"J","Q","K"),4),c("H","D","S","C"))
 sample(cards,5)               # a pair of jacks, no replacement
```
```{r}
## roll 2 die. Even fancier
 dice = as.vector(outer(1:6,1:6,paste))
 sample(dice,5,replace=TRUE)   # replace when rolling dice
```
The last two illustrate things that can be done with a little typing and a lot of thinking using the fun commands paste for pasting together strings, rep for repeating things and outer for generating all possible products.


2.3A bootstrap sample
Bootstrapping is a method of sampling from a data set to make statistical inference. The intuitive idea is that by sampling, one can get an idea of the variability in the data. The process involves repeatedly selecting samples and then forming a statistic. Here is a simple illustration on obtaining a sample.

The examples in this section are based on a built-in data frame named faithful. It consists of a collection of observations of the Old Faithful geyser in the USA Yellowstone National Park. The following is a preview via the head function.
```{r}
head(faithful)
```

There are two observation variables in the data set. The first one, called eruptions, is the duration of the geyser eruptions. The second one, called waiting, is the length of waiting period until the next eruption.

The built in data set faithful has a variable "eruptions'' that measures the time between eruptions at Old Faithful. It has an unusual distribution. A bootstrap sample is just a sample with replacement from the given values. It can be found as follows

```{r}
data(faithful)                # part of R's base
 names(faithful)               # find the names for faithful
```
```{r}
eruptions = faithful[['eruptions']] # or attach and detach faithful
 sample(eruptions,10,replace=TRUE)
```
```{r}
hist(eruptions,breaks=25)     # the dataset
```
```{r}
## the bootstrap sample
 hist(sample(eruptions,100,replace=TRUE),breaks=25)
```

2.4 d,p and q functions
The d functions were used to plot the theoretical densities above. As with the "r" functions, you need to specify the parameters, but differently, you need to specify the x values (not the number of random numbers n).

The p and q functions are for the cumulative distribution functions and the quantiles. As mentioned, the distribution of a random number is specified by the probability that the number is between a and b for arbitrary a and b,P(a<X<??????b). In fact, the value F(x) =P(X<=??????b) is enough.

The p functions answer what is the probability that a random variable is less than x. Such as for a standard normal, what is the probability it is less than .7?
```{r}
pnorm(.7)                     # standard normal
```
```{r}
pnorm(.7,1,1)                 # normal mean 1, std 1
```
Notationally, these answer P(Z<=???.7) where Z is a standard normal variable or normal(1,1). To answer P(Z> .7) is also easy. You can do the work by noting this is 1 -P(Z<=.7) or let R do the work, by specifying lower.tail=F as in:

```{r}
pnorm(.7,lower.tail=F)
```
The q function are inverse to this. They ask, what value corresponds to a given probability. This the quantile or point in the data that splits it accordingly. For example, what value of z has .75 of the area to the right for a standard normal? (This is Q3)

```{r}
qnorm(.75)
```
Notationally, this is finding z which solves 0.75 =P(Z<=z).


2.5 Standardizing,scale and z scores
To standardize a random variable you subtract the mean and then divide by the standard deviation. That is
                            Z=(X-??)/??
                            
To do so requires knowledge of the mean and standard deviation.

You can also standardize a sample. There is a convenient function scale that will do this for you. This will make your sample have mean 0 and standard deviation 1. This is useful for comparing random variables which live on different scales.

Normal random variables are often standardized as the distribution of the standardized normal variable is again normal with mean 0 and variance 1. (The standard'' normal.) The z-score of a normal number is the value of it after standardizing.

If we have normal data with mean 100 and standard deviation 16 then the following will find the z-scores
```{r}
x = rnorm(5,100,16)
x
```
```{r}
z = (x-100)/16
z
```
The z-score is used to look up the probability of being to the right of the value of x for the given random variable. This way only one table of normal numbers is needed. With R, this is not necessary. We can use the pnorm function directly
```{r}
pnorm(z)
```
```{r}
pnorm(x,100,16)               # enter in parameters
```






                            
                            
                            
          Exercise 2.1
              Generate 10 random numbers from a uniform distribution on [0,10].                Use R to find the maximum and minimum values.x
```{r}
                    num=runif(10,0,10)
                    print(num)
```
To find the maximum and the minimum values we must
have a number to assure that. let's say in range 
[0,10] that this number is:
```{r}
                      med=(0+10)/2
                      print(med)
```
Let's find the minimum and maximum values:

```{r}
                    print("The maximum values:")
                    for (val in num){
                      if(val>med){
                        print(val)
                        
                      }
                    }



                   print("The minimum values:")
                    for (val in num){
                      if(val<med){
                        print(val)
                        
                      }
                    }

```
                    
          
          
          
          Exercise 2.2
          Generate 10 random normal numbers with mean 5 and standard deviation 5           (normal(5,5)). How many are less than 0? (Use R)

          Counting
          Frequently  it  is  useful  to  count  how  many  elements  within  a            vector  satisfy some condition.  For example, if we wanted to know how           many of the integers between  1  and 100  are  divisible by  7,  we              could use
```{r}
          A = seq(100)
          B = (A %% 7 == 0)
          x = sum(B)
          print(x)
```

        Note that when summing the Boolean vector B, true values are counted as 1
        and  false values  are  counted as  0.
        
```{r}
        randnums=rnorm(10,5,5)
        bool=(randnums<0)
        ans=sum(bool)
        print("The random normal nums :")
        print(randnums)
        print("The # of the nums that are less than 0 are:")
        print(ans)
```
        
          Exercise 2.3
          Generate 100 random normal numbers with mean 100 and standard                   deviation 10. How many are 2 standard deviations
          from the mean (smaller than 80 or bigger than 120)?
          
```{r}
        m=100
        s=10
        x=rnorm(100,m,s)
        q1=(x<m-s)
        sm=sum(q1)
        q2=(x>m+s)
        bi=sum(q2)
        smorbi=sm+bi
        print(smorbi)
```
          
          Exercise 2.4
          Toss a fair coin 50 times (using R). How many heads do you have?
          
```{r}
        coin<- c('H','T')
        Toss=function(n) sample(coin,n,replace=T)
        q3=Toss(50)
        print("Here are the 50 tosses:")
        print(q3)
        heads=sum(q3=="H")
        print("The # of heads:")
        print(heads)
```
          
          Exercise 2.5
          Roll a die'' 100 times. How many 6's did you see?
          
```{r}
          RollDie=function(n) sample(1:6,n,replace=T)
          q4=RollDie(100)
          print(q4)
          sixes=sum(q4==6)
          print('The 6 s we have are:')
          print(sixes)
          
```
          
          Exercise 2.6
          Select 6 numbers from a lottery containing 49 balls. What is the                largest number? What is the smallest? Answer these using R.
          
          
```{r}
          Lottery=function(n) sample(1:49,n)
          selection=Lottery(6)
          print('The selected nums from the lottery:')
          print(selection)
          print('The largest number:')
          print(max(selection))
           print('The smallest number:')
          print(min(selection))
          
```
          
          
          Exercise 2.7
          For normal(0,1), find a number z* solving P(Z<=???z*) = .05 (use qnorm)
          
```{r}
          z1=qnorm(0.05)
          print(z1)
```
          
          Exercise 2.8
          For normal(0,1), find a number z* solving P(-z*<=???<=Z??????z*) = .05                 (use qnorm and symmetry)
          
          We see that P(-z*<=Z<=z*)=P(Z<=z*)-P(Z<=-z*)
```{r}
          z2=qnorm(0.05/2)
          print(z2)

```
          
          Exercise 2.9
          How much area (probability) is to the right of 1.5 for a normal(0,2)?
          
```{r}
          p=pnorm(1.5,0,2,lower.tail=F)
          print(p)
```
        
          Exercise 2.10
          Make a histogram of 100 exponential numbers with mean 10. Estimate the           median. Is it more or less than the mean?
          
          
```{r}
          expo=rexp(100,1/10)
          hist(expo)
```
          The median must be 20,...
          Let's see if our estimation is right:
```{r}
          median(expo)
```
         We see after all that the median is smaller from the mean 
          
          
          
          Exercise 2.11
          Can you figure out what this R command does?

          > rnorm(5,mean=0,sd=1:5)
          
          My estimation is that it creates 5 different normal distributions each           with different standard deviation from 1 to 5 and consisted of 5                random normal numbers
          
```{r}
          rnorm(5,mean=0,sd=1:5)
```
            Wrong!! It seems that it creates 5 random normal numbers,each with 
            the same mean but with different standard deviation,counting from
            1 to 5
          
          
          
          
          Exercise 2.12