---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Exercise 1

Suppose the manufacturer claims that the mean lifetime of a light bulb is more than 10,000
hours. In a sample of 30 light bulbs, it was found that they only last 9,900 hours on average.
Assume the population standard deviation is 120 hours. At .05 significance level, can we reject
the claim by the manufacturer?
```{r}
X=10000
m=9900
sd=120
n=30
sub_z1=X-m
sub_z2=sd/sqrt(n)
z=sub_z1/sub_z2
z
a=0.95
z_a=qnorm(1-a)
z_a
```

We must first define our null hypothesis.Our null hypothesis is obviously the claim
of the manufacturer.That is,that the mean lifetime of a light-bulb is more than
10000 hours.We,now,see that at significance level 0.05 z>-z_a is true,so we are not going to reject the null hypothesis,so as the claim by the manufacturer.

Exercise 2

Suppose the food label on a cookie bag states that there is at most 2 grams of saturated fat in a
single cookie. In a sample of 35 cookies, it is found that the mean amount of saturated fat per
cookie is 2.1 grams. Assume that the population standard deviation is 0.25 grams. At .05
significance level, can we reject the claim on food label?
```{r}
X=2
m=2.1
sd=0.25
n=35
sub_z1=X-m
sub_z2=sd/sqrt(n)
z=sub_z1/sub_z2
z
a=0.95
z_a=qnorm(1-a)
z_a
```
The null hypothesis here is that there is at most 2 grams of saturated fat in a
simple cookie.With a significance level of 0.05 we see that z<z_a.That means that
the null hypothesis is not rejected,so we don't reject the claim on the food label

Exercise 3

Suppose the mean weight of King Penguins found in an Antarctic colony last year was 15.4 kg.
In a sample of 35 penguins same time this year in the same colony, the mean penguin weight is
14.6 kg. Assume the population standard deviation is 2.5 kg. At .05 significance level, can we
reject the null hypothesis that the mean penguin weight does not differ from last year?
For the computation of critical value we use the qnorm(1???alpha/2). Why?

```{r}
X=15.4
m=14.6
sd=2.5
n=35
sub_z1=X-m
sub_z2=sd/sqrt(n)
z=sub_z1/sub_z2
z
a=0.95
a=a/2
z_a=qnorm(1-a)
z_a
```
We see that z>z_a/2, so we reject the null hypothesis that  the mean penguin weight does not differ from last year

 In order to reach the hypothesis m=m_0
we must include the hypothesis m<m_0, and m>m_0 each with significance level
a=0.05.So that's why we use a/2 for our first hypothesis.


Exercise 4

Suppose the manufacturer claims that the mean lifetime of a light bulb is more than 10,000
hours. In a sample of 30 light bulbs, it was found that they only last 9,900 hours on average.
Assume the sample standard deviation is 125 hours. At .05 significance level, can we reject the
claim by the manufacturer?
```{r}
X=10000
m=9900
sd=125
n=30
sub_t1=X-m
sub_t2=sd/sqrt(n)
t=sub_t1/sub_t2
t
t_a=qt(p=0.025,df=29)
t_a
```
We see that t>-t_a so we don't reject the claim

Exercise 5

Suppose the food label on a cookie bag states that there is at most 2 grams of saturated fat in a
single cookie. In a sample of 35 cookies, it is found that the mean amount of saturated fat per
cookie is 2.1 grams. Assume that the sample standard deviation is 0.3 gram. At .05 significance
level, can we reject the claim on food label?
```{r}
X=2
m=2.1
sd=0.3
n=35
sub_t1=X-m
sub_t2=sd/sqrt(n)
t=sub_t1/sub_t2
t
a=0.95
t_a=qt(p=0.025,df=34)
t_a
```
Indeed,we have t>t_a so we reject the claim

Exercise 6
Suppose the mean weight of King Penguins found in an Antarctic colony last year was 15.4 kg.
In a sample of 35 penguins same time this year in the same colony, the mean penguin weight is
14.6 kg. Assume the sample standard deviation is 2.5 kg. At .05 significance level, can we reject
the null hypothesis that the mean penguin weight does not differ from last year?
```{r}
X=15.4
m=14.6
sd=2.5
n=35
sub_t1=X-m
sub_t2=sd/sqrt(n)
t=sub_t1/sub_t2
t
t_a=qt(p=0.025,df=34)
t_a
```
t>t_a and t<-t_a so the null hypothesis is going to be rejected

Exercise 7
Suppose 60% of citizens voted in last election. 85 out of 148 people in a telephone survey said
that they voted in current election. At 0.5 significance level, can we reject the null hypothesis
that the proportion of voters in the population is above 60% this year?
```{r}
p=85/148
p_0=0.6
subset_z1=p-p_0
z2=1-p_0
z2=p_0*z2
z2=z2/148
subset_z2=sqrt(z2)
z=subset_z1/subset_z2
z
z_a=qnorm(0.5)
z_a
```
We have z<z_a so we keep the null hypothesis

Exercise 8
Suppose that 12% of apples harvested in an orchard last year was rotten. 30 out of 214 apples in
a harvest sample this year turns out to be rotten. At .05 significance level, can we reject the null
hypothesis that the proportion of rotten apples in harvest stays below 12% this year?
```{r}
p=30/214
p_0=0.12
subset_z1=p-p_0
z2=1-p_0
z2=p_0*z2
z2=z2/214
subset_z2=sqrt(z2)
z=subset_z1/subset_z2
z
z_a=qnorm(0.95)
z_a
```

z<z_a,so we keep the null hypothesis

Exercise 9

Suppose a coin toss turns up 12 heads out of 20 trials. At .05 significance level, can one reject
the null hypothesis that the coin toss is fair?

```{r}
p=12/20
p_0=0.5
subset_z1=p-p_0
z2=1-p_0
z2=p_0*z2
z2=z2/20
subset_z2=sqrt(z2)
z=subset_z1/subset_z2
z
z_a=qnorm(0.95)
z_a
```
-z_a<z<z_a so we can't reject the null hypothesis

```{r}
mean=c()
table=c(10, 2, 1, 1, 15, 7, 4, 18)
for(i in table){
  for(j in table){
     if(i==j){
       next
     }
     x=(i+j)/2
     mean<-append(mean,x)
  }
  
  
}
   
    
hist(mean,breaks=seq(0,18,0.5))
mean
```







